### 2017-07-05 README.Rmd
update ReadMe

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# WetlandComparison

The goal of WetlandComparison is to make a first attempt at a reproducible research R package. This will still be somewhat hampered by the fact that much of this research is conducted with Matlab.

## Installation

You can install WetlandComparison from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("samdchamberlain/WetlandComparison")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```