### 2018-02-23 README.Rmd
Readme update

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![Travis-CI Build Status](https://travis-ci.org/samdchamberlain/wetlandcomparison.svg?branch=master)](https://travis-ci.org/samdchamberlain/wetlandcomparison)

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# WetlandComparison

This package was used the generate the manuscript "Soil properties and sediment accretion modulate methane fluxes from restored wetlands." by Samuel D. Chamberlain et. al. This follows the R package structure adapted from Carl Boettiger's template found here: https://github.com/cboettig/template. Briefly, the /data folder contains soil and half-hourly eddy covariance files used in the analysis, the /R folder contains scripts used to process data, and the /vignettes folder contains the Markdown and bibiography files used to generate the manuscript, as well as two figure images that were not directly reproducible in R.

Figure 1 and 4 are not directly reproducible in R, but all other analyses are reproducible in this package. Figure 1 was generated in ArcMap using the soil series data downloaded from the USDA Web Soil Survey (https://websoilsurvey.sc.egov.usda.gov/) and eddy covariance flux footprints. Figure 4 was generated using a combined information theory-wavelet time series decomposition using a modification of the MATLAB ProcessNetwork Software found here: https://github.com/samdchamberlain/ProcessNetwork_Software. All other analyses in this work are reproducible from the 'manuscript.Rmd' file in the /vignettes folder, and running this file will reproduce this manuscript in its entirety.

## Installation

You can install WetlandComparison from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("samdchamberlain/WetlandComparison")
```

## How to Run
Load package into RStudio and open the 'manuscript.Rmd' within the /vignettes folder. Using RStudio, click the 'Knit' button at the top of the console and this manuscript, and analyses herein, will be re-created.